Single-cell virus sequencing of influenza variants that trigger innate immunity

### 2018-09-03 paper/paper.tex
added plot on ISG / virus associations to paper
Added figure supplement to Figure 4 that shows
the association of viral gene absence / mutations
with ISG expression. Trends are mostly similar
as for IFN expression.

%%% begin mutations figure %%%
\begin{figure}
\begin{fullwidth}
{\centering
\includegraphics[width=\linewidth]{figures/single_cell_figures/p_mutations.pdf}
}
\caption{
Viral genetic variation partially explains the heterogeneity in viral burden and IFN induction among the infected cells for which we sequenced all expressed viral genes (\FIG{genotypes}).
{\bf (A)} 
The percent of all mRNA derived from virus, faceted by whether the cells express wildtype copies of all eight viral genes, or fail to express a wildtype copy of one or more genes (due to gene absence, amino-acid substitution, insertion, or deletion).
Cells infected by full wildtype virus exhibit less heterogeneity in viral burden as quantified by the Gini index (95\% confidence intervals are indicated).
{\bf (B)}
There is a trend for cells infected by full wildtype virus to express less IFN.
{\bf (C)}
Some specific forms of viral genetic variation are associated with IFN induction.
The top panel show the percent of IFN+ and IFN- cells that fail to express each viral gene.
The middle and bottom panels show the percent of IFN- and IFN+ cells that have a deletion or amino-acid substitution in each viral gene, conditioned on the cell expressing that gene.
Numbers above the bars give P-values (Fisher's exact test) and associated Q-values~\citep{storey2003statistical} for rejecting the null hypothesis that percents are equal among IFN- and IFN+ cells. 
Absence of NS is strongly associated with IFN induction, and amino-acid substitutions in PB1 or NS and deletions in HA are weakly associated with IFN induction.
Only the absence of NS remains significant at a false-discovery rate of 10\%.
\FIGSUPP[mutations]{allcells} shows that the trends remain largely the same if we also analyze infected cells with incomplete viral sequence information, with the addition of a trend towards IFN induction associated with \emph{presence} of PB2 and PA. 
Insertions are not shown as a mutation type as they are extremely rare (\FIG{genotypes}).
{\bf (D)}
There is no association between IFN induction and the amount of viral mRNA in cells that express NS, but viral burden is associated with IFN induction among cells that lack NS.
Note that throughout this figure, we only consider as substitution mutations that are non-synonymous.
}
\label{fig:mutations}

\figsupp[Analysis as in \FIG{mutations}C but including infected cells with incomplete viral sequence information.]
{This plot differs from \FIG{mutations}C in that it also includes data from cells for which some viral genes were not fully sequenced.
For incompletely sequenced cells, deletions and substitutions are included in the counts when that particular viral gene is sequenced.
The major trends in \FIG{mutations}C are also true for the larger dataset in this figure.
In addition, there is now a modest trend for infected cells that fail to express PB2 and PA \emph{not} to express IFN.
}
{\includegraphics[width=\textwidth]{figures/single_cell_figures/p_muts_ifn_all_annot.pdf}}
\label{figsupp:allcells}

>>>>

%%% begin mutations figure %%%
\begin{figure}
\begin{fullwidth}
{\centering
\includegraphics[width=\linewidth]{figures/single_cell_figures/p_mutations.pdf}
}
\caption{
Viral genetic variation partially explains the heterogeneity in viral burden and IFN induction among the infected cells for which we sequenced all expressed viral genes (\FIG{genotypes}).
{\bf (A)} 
The percent of all mRNA derived from virus, faceted by whether the cells express wildtype copies of all eight viral genes, or fail to express a wildtype copy of one or more genes (due to gene absence, amino-acid substitution, insertion, or deletion).
Cells infected by full wildtype virus exhibit less heterogeneity in viral burden as quantified by the Gini index (95\% confidence intervals are indicated).
{\bf (B)}
There is a trend for cells infected by full wildtype virus to express less IFN.
{\bf (C)}
Some specific forms of viral genetic variation are associated with IFN induction.
The top panel show the percent of IFN+ and IFN- cells that fail to express each viral gene.
The middle and bottom panels show the percent of IFN- and IFN+ cells that have a deletion or amino-acid substitution in each viral gene, conditioned on the cell expressing that gene.
Numbers above the bars give P-values (Fisher's exact test) and associated Q-values~\citep{storey2003statistical} for rejecting the null hypothesis that percents are equal among IFN- and IFN+ cells. 
Absence of NS is strongly associated with IFN induction, and amino-acid substitutions in PB1 or NS and deletions in HA are weakly associated with IFN induction.
Only the absence of NS remains significant at a false-discovery rate of 10\%.
\FIGSUPP[mutations]{allcells} shows that the trends remain largely the same if we also analyze infected cells with incomplete viral sequence information, with the addition of a trend towards IFN induction associated with \emph{presence} of PB2 and PA. 
\FIGSUPP[mutations]{isg} shows that the trends also remain largely similar if we analyze the association of viral genetic variation with ISG expression rather than IFN expression.
Insertions are not shown as a mutation type as they are extremely rare (\FIG{genotypes}).
{\bf (D)}
There is no association between IFN induction and the amount of viral mRNA in cells that express NS, but viral burden is associated with IFN induction among cells that lack NS.
Note that throughout this figure, we only consider as substitution mutations that are non-synonymous.
}
\label{fig:mutations}

\figsupp[Analysis as in \FIG{mutations}C but including infected cells with incomplete viral sequence information.]
{This plot differs from \FIG{mutations}C in that it also includes data from cells for which some viral genes were not fully sequenced.
For incompletely sequenced cells, deletions and substitutions are included in the counts when that particular viral gene is sequenced.
The major trends in \FIG{mutations}C are also true for the larger dataset in this figure.
In addition, there is now a modest trend for infected cells that fail to express PB2 and PA \emph{not} to express IFN.
}
{\includegraphics[width=\textwidth]{figures/single_cell_figures/p_muts_ifn_all.pdf}}
\label{figsupp:allcells}

\figsupp[Association of viral genetic variation and ISG expression.]
{This plot shows the same information as panels \FIG{mutations}B-D except that it shows the association of viral gene absence / mutations with ISG expression rather than IFN expression.
Cells are classified as ISG+ or ISG- as in \FIGSUPP[experiment]{ISG}.
The major viral features that associate with IFN induction also associate with ISG expression.
Specifically, cells that express wildtype copies of all eight viral genes tend to express ISGs less frequently and at lower levels
The absence of NS is strongly associated with higher ISG expression.
Viral burden is not associated with ISG expression levels in NS-sufficient infections, but is in NS-deficient infections.
}
{\includegraphics[width=\textwidth]{figures/single_cell_figures/p_mutations_isg.pdf}}
\label{figsupp:isg}
